NFL Draft Analysis

### 2020-05-12 draft_analysis.Rmd
Updated prose to add more details

Next we will scrape the AP Offensive Rookie of the Year Winners from the same website, again for our machine learning problem.

@@

Load the NFL draft history data from *master_df*, containing the draft data from rounds 1-7 from the 1970-2016 NFL draft. Remove all of the missing data entries, and convert the necessary columns to integers.

@@

Now let's process our 2020 Draft data. First, we will get rid of columns we don't care about, then remove blank rows, add year drafted to each observation, and convert the necessary columns to integers, all like we did for the historical draft dataset. We will also filter this dataframe to only include quarterbacks, wide receivers, and running backs.

@@

We will process the *test_df* dataframe in the same way.

@@

Has average value for players selected in different rounds gone up over time?
Have GMs gotten better at drafting in the first round, second round, etc...?

@@

@@

Now let's focus our attention on the past 20 years and graph some visualizations of that data.
```{r qb2}

@@

@@

Next we will provide the same graphs to provide similar insight into the RB and WR selections of the draft.

@@

Let's test our model against the 2017, 2018, and 2019 datasets. To do this, we will use the *predict()* function in the caret package and feed it the test data. From there,  *predict()* will calculate probabilities that the model is sure about it choice that each observation from *test_df* will fall under the "Yes" and "No" outcome categories. So the "Yes" variable corresponds to the probability the model is sure about its choice that the player will become Rookie MVP, and "No" = probability the model is sure the player will not be the Rookie MVP.

@@

Now let's join our prediction probabilities to the actual data, so we can look at the prediction probabilities for each player.

@@

Now let's join our prediction probabilities to the actual data like we did before. s

@@

Joe Burrow is our model's top prediction to win the MVP!!! We also see that Cam Akers, Tua, Henry Ruggs III, Jerry Jeudy, CeDee Lamb, and Justin Jefferson also appear in the top 10. All of these players are among those that have the highest odds to win the the Offensive MVP according to OddsShark.

>>>>

Next we will scrape the AP Offensive Rookie of the Year Winners from the same website, again for our machine learning problem, and save into the *roty_df* dataframe.

@@

Load the NFL draft history data from *master_df*, containing the draft data from rounds 1-7 from the 1970-2016 NFL draft. We will only keep the observations that do not have any missing data entries. Also, we will convert the necessary columns to integers and factors (rather than chars). This way, we can use the columns values to group and to produce regression lines in our data visualization steps later.

@@

Now let's process our 2020 Draft data. First, we will get rid of columns we don't care about, then remove blank rows, add year drafted to each observation, and convert the necessary columns to integers and factors, all like we did for the historical draft dataset. We will also filter this dataframe to only include quarterbacks, wide receivers, and running backs.

@@

We will process the *test_df* dataframe in the same way as the *df_2020* dataframe.

@@

Our first visualization will answer the following questions for us:
Has average value for players selected in different rounds gone up over time?
Have GMs gotten better at drafting in the first round, second round, etc...?

To do this, we will plot the average value for all the players for each year since 1970, grouped by round.

@@

So, lets create a bar graph. For each time period, we will graph the average value of a quarterback selected in each round.

@@

Now let's focus our attention on the past 20 years and graph some visualizations of that data.

Our first graph for the 2000s will be the average value of a quarterback on the y-axis, and the pick number in which they were selected on the x-axis.

```{r qb2}

@@

Our next plot will we a violin plot of the average value for quarterbakcs drafted in the 2000s in different rounds. This will give us a better idea of the distribution of the average value for the quarterback position.

@@

Next we will provide the same graphs to provide similar insight into the RB and WR selections of the draft. We will use *grid.arrange* to show the graphs side by side in three columns.

@@

Let's test our model against the 2017, 2018, and 2019 datasets. To do this, we will use the *predict()* function in the caret package and feed it the test data. From there,  *predict()* will calculate probabilities that the model is sure about it choice that each observation from *test_df* will fall under the "Yes" and "No" outcome categories. So the **Yes** variable corresponds to the probability the model is sure about its choice that the player will become Rookie MVP, and **No** = probability the model is sure the player will not be the Rookie MVP.

@@

Now let's join our prediction probabilities to the actual data, so we can look at the prediction probabilities for each player. We will use *merge*, which concatentaes dataframes horizontally. Setting the "by" parameter to "row.names" matches the dataframes by row index.

@@

Now let's join our prediction probabilities to the actual data like we did before.

@@

**Joe Burrow is our model's top prediction to win the MVP!!!** We also see that Cam Akers, Tua, Henry Ruggs III, Jerry Jeudy, CeDee Lamb, and Justin Jefferson also appear in the top 10. All of these players are among those that have the highest odds to win the the Offensive MVP according to OddsShark.
